---
title: "Factorization methods for solving Ax = b"
authors: Sofiya Romanchuk, Anna Pynda
---

```{r}
library(QuantumOps)
library(matrixcalc)
library(matlib)
```

QR factorization via Gram-Schmidt algorithm.
```{r}
QR_own <- function(x){
  
  n <- nrow(x)
  m <- ncol(x)
  
  U <- matrix(0, nrow = n, ncol = m)
  Q <- matrix(0, nrow = n, ncol = m)
  R <- matrix(0, nrow = m, ncol = m) 
  
  U[, 1] <- x[, 1]
  
  for (k in 2:ncol(x)) {
    U[, k] <- x[, k]
    #Creating orthogonal system of vectors using Gram-Schmidt algorithm
    for (w in seq(1, (k - 1), 1)) {
      U[, k] <- U[, k] - ((inner(U[, w], x[, k]) / inner(U[, w], U[, w])) * (U[, w]))
    }
  }
  
  #Normalizing orthogonal system
  normalized <- function(x) {
    x / sqrt(inner(x, x)) 
  }
  #Final calculations of Q and R
  Q <- apply(U, 2, normalized)
  R <- crossprod(Q, x)
  
  return(list(Q = Q, R = R))
}
```

LU factorization via Doolitle algoritm.
```{r}
LU_own <- function(x){
    n <- nrow(x)
    n <- ncol(x)
    L <- matrix(0, nrow = n, ncol = n)
    U <- matrix(0, nrow = n, ncol = n)
    
    diag(L) <- rep(1, n)
    #Decomposing matrix into Upper- and Lower-triangular matrix 
    for (i in 1:n) {
        u <- i + 1
        l <- i - 1
        #Determining Upper-triangular matrix
        for (j in 1:n) {
            U[i,j] <- x[i,j]
            if (l > 0) {
                for (k in 1:l) {
                  #Evaluating U[i, j] via summation of L[i,k] * U[k,j]
                    U[i,j] <- U[i,j] - L[i,k]*U[k,j]
                }
            }
        }
        #Determining Lower-triangular matrix
        if (u <= n) {
            for (j in u:n) {
                L[j,i] <- x[j,i]
                if ( l > 0 ) {
                    for (k in 1:l) {
                      #Evaluating L[j, i] via summation of L[j,k] * U[k,i]
                        L[j,i] <- L[j,i] - L[j,k]*U[k,i]
                    }
                }
                if (U[i,i] == 0)
                    stop()
                L[j,i] <- L[j,i]/U[i,i]
            }    
        }
    }
    return(list(L=L, U=U))
}
```

Time tests for LU, QR and Cholesky factorizations, in comparsion via built-in functions and our own implementation of algorithms, on different matrices

Part that includes time tests on symmetric non-singular matrices for Cholesky, LU and QR factorizations.
#timing factorization methods on the 3x3 matrices
```{r}
sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(25, 15, -5, 15, 18, 0, -5, 0, 11), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  lu.decomposition(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by3 matrix LU own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(25, 15, -5, 15, 18, 0, -5, 0, 11), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  LU_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by3 matrix LU")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(25, 15, -5, 15, 18, 0, -5, 0, 11), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  QR(m1)
  time_lu = Sys.time() - t0
  
  sum1 = sum1 + time_lu
}
print("3by3 matrix QR")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(25, 15, -5, 15, 18, 0, -5, 0, 11), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  QR_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by3 matrix QR own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(25, 15, -5, 15, 18, 0, -5, 0, 11), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  cholesky(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by3 matrix Cholesky")
print(sum1/1000)
```

```{r}
sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(6, 15, 55, 15, 55, 225, 55, 225, 979), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  lu.decomposition(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by3 matrix LU own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(6, 15, 55, 15, 55, 225, 55, 225, 979), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  LU_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by3 matrix LU")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(6, 15, 55, 15, 55, 225, 55, 225, 979), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  QR(m1)
  time_lu = Sys.time() - t0
  
  sum1 = sum1 + time_lu
}
print("3by3 matrix QR")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(6, 15, 55, 15, 55, 225, 55, 225, 979), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  QR_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by3 matrix QR own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(6, 15, 55, 15, 55, 225, 55, 225, 979), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  cholesky(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by3 matrix Cholesky")
print(sum1/1000)
```

```{r}
sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(6, -2, 2, -2, 3, -1, 2, -1, 3), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  lu.decomposition(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by3 matrix LU own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(6, -2, 2, -2, 3, -1, 2, -1, 3), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  LU_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by3 matrix LU")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(6, -2, 2, -2, 3, -1, 2, -1, 3), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  QR(m1)
  time_lu = Sys.time() - t0
  
  sum1 = sum1 + time_lu
}
print("3by3 matrix QR")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(6, -2, 2, -2, 3, -1, 2, -1, 3), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  QR_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by3 matrix QR own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(6, -2, 2, -2, 3, -1, 2, -1, 3), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  cholesky(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by3 matrix Cholesky")
print(sum1/1000)
```

```{r}
sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(25, 15, -5, 15, 18, 0, -5, 0, 11), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  lu.decomposition(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by3 matrix LU own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(25, 15, -5, 15, 18, 0, -5, 0, 11), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  LU_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by3 matrix LU")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(25, 15, -5, 15, 18, 0, -5, 0, 11), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  QR(m1)
  time_lu = Sys.time() - t0
  
  sum1 = sum1 + time_lu
}
print("3by3 matrix QR")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(25, 15, -5, 15, 18, 0, -5, 0, 11), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  QR_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by3 matrix QR own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(25, 15, -5, 15, 18, 0, -5, 0, 11), nrow=3, byrow=TRUE )
  t0 = Sys.time()
  cholesky(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by3 matrix Cholesky")
print(sum1/1000)
```

#timing factorization methods on the 4x4 matrices
```{r}
sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(16, -12, -12, -16, -12, 25, 1, -4, -12, 1, 17, 14, -16, -4, 14, 57), nrow=4, byrow=TRUE )
  t0 = Sys.time()
  lu.decomposition(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("4by4 matrix LU own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(16, -12, -12, -16, -12, 25, 1, -4, -12, 1, 17, 14, -16, -4, 14, 57), nrow=4, byrow=TRUE )
  t0 = Sys.time()
  LU_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("4by4 matrix LU")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(16, -12, -12, -16, -12, 25, 1, -4, -12, 1, 17, 14, -16, -4, 14, 57), nrow=4, byrow=TRUE )
  t0 = Sys.time()
  QR(m1)
  time_lu = Sys.time() - t0
  
  sum1 = sum1 + time_lu
}
print("4by4 matrix QR")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(16, -12, -12, -16, -12, 25, 1, -4, -12, 1, 17, 14, -16, -4, 14, 57), nrow=4, byrow=TRUE )
  t0 = Sys.time()
  QR_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("4by4 matrix QR own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(16, -12, -12, -16, -12, 25, 1, -4, -12, 1, 17, 14, -16, -4, 14, 57), nrow=4, byrow=TRUE )
  t0 = Sys.time()
  cholesky(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("4by4 matrix Cholesky")
print(sum1/1000)
```

```{r}
sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(9, 0, -27, 18, 0, 9, -9, -27, -27, -9, 99, -27, 18, -27, -27, 121), nrow=4, byrow=TRUE )
  t0 = Sys.time()
  lu.decomposition(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("4by4 matrix LU own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(9, 0, -27, 18, 0, 9, -9, -27, -27, -9, 99, -27, 18, -27, -27, 121), nrow=4, byrow=TRUE )
  t0 = Sys.time()
  LU_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("4by4 matrix LU")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(9, 0, -27, 18, 0, 9, -9, -27, -27, -9, 99, -27, 18, -27, -27, 121), nrow=4, byrow=TRUE )
  t0 = Sys.time()
  QR(m1)
  time_lu = Sys.time() - t0
  
  sum1 = sum1 + time_lu
}
print("4by4 matrix QR")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(9, 0, -27, 18, 0, 9, -9, -27, -27, -9, 99, -27, 18, -27, -27, 121), nrow=4, byrow=TRUE )
  t0 = Sys.time()
  QR_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("4by4 matrix QR own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(9, 0, -27, 18, 0, 9, -9, -27, -27, -9, 99, -27, 18, -27, -27, 121), nrow=4, byrow=TRUE )
  t0 = Sys.time()
  cholesky(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("4by4 matrix Cholesky")
print(sum1/1000)
```

```{r}
sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(10, -1, 2, 0, -1, 11, -1, 3, 2, -1, 10, -1, 0, 3, -1, 8), nrow=4, byrow=TRUE )
  t0 = Sys.time()
  lu.decomposition(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("4by4 matrix LU own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(10, -1, 2, 0, -1, 11, -1, 3, 2, -1, 10, -1, 0, 3, -1, 8), nrow=4, byrow=TRUE )
  t0 = Sys.time()
  LU_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("4by4 matrix LU")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(10, -1, 2, 0, -1, 11, -1, 3, 2, -1, 10, -1, 0, 3, -1, 8), nrow=4, byrow=TRUE )
  t0 = Sys.time()
  QR(m1)
  time_lu = Sys.time() - t0
  
  sum1 = sum1 + time_lu
}
print("4by4 matrix QR")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(10, -1, 2, 0, -1, 11, -1, 3, 2, -1, 10, -1, 0, 3, -1, 8), nrow=4, byrow=TRUE )
  t0 = Sys.time()
  QR_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("4by4 matrix QR own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(10, -1, 2, 0, -1, 11, -1, 3, 2, -1, 10, -1, 0, 3, -1, 8), nrow=4, byrow=TRUE )
  t0 = Sys.time()
  cholesky(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("4by4 matrix Cholesky")
print(sum1/1000)
```

#timing factorization methods on the 6x6 matrices
```{r}
sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(7, -2, -3, 0, -1, 0, -2, 8, 0, 0, -1, 0, -3, 0, 4, -1, 0, 0, 0, 0, -1, 5, 0, -2, -1, -1, 0, 0, 4, 0, 0, 0, 0, -2, 0, 6), nrow=6, byrow=TRUE )
  t0 = Sys.time()
  lu.decomposition(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("6by6 matrix LU own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(7, -2, -3, 0, -1, 0, -2, 8, 0, 0, -1, 0, -3, 0, 4, -1, 0, 0, 0, 0, -1, 5, 0, -2, -1, -1, 0, 0, 4, 0, 0, 0, 0, -2, 0, 6), nrow=6, byrow=TRUE )
  t0 = Sys.time()
  LU_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("6by6 matrix LU")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(7, -2, -3, 0, -1, 0, -2, 8, 0, 0, -1, 0, -3, 0, 4, -1, 0, 0, 0, 0, -1, 5, 0, -2, -1, -1, 0, 0, 4, 0, 0, 0, 0, -2, 0, 6), nrow=6, byrow=TRUE )
  t0 = Sys.time()
  QR(m1)
  time_lu = Sys.time() - t0
  
  sum1 = sum1 + time_lu
}
print("6by6 matrix QR")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(7, -2, -3, 0, -1, 0, -2, 8, 0, 0, -1, 0, -3, 0, 4, -1, 0, 0, 0, 0, -1, 5, 0, -2, -1, -1, 0, 0, 4, 0, 0, 0, 0, -2, 0, 6), nrow=6, byrow=TRUE )
  t0 = Sys.time()
  QR_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("6by6 matrix QR own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c(7, -2, -3, 0, -1, 0, -2, 8, 0, 0, -1, 0, -3, 0, 4, -1, 0, 0, 0, 0, -1, 5, 0, -2, -1, -1, 0, 0, 4, 0, 0, 0, 0, -2, 0, 6), nrow=6, byrow=TRUE )
  t0 = Sys.time()
  cholesky(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("6by6 matrix Cholesky")
print(sum1/1000)
```

Part that includes time tests on square matrices for QR and LU decompositions.
#timing factorization methods on the 2x2 matrices
```{r}
sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c ( 1, 2, 3, 4 ), nrow=2, byrow=TRUE)
  t0 = Sys.time()
  LU_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("2by2 matrix LU")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c ( 1, 2, 3, 4 ), nrow=2, byrow=TRUE)
  t0 = Sys.time()
  lu.decomposition(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("2by2 matrix LU own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix(c(2, 1, 1, 2), ncol=2)
  t0 = Sys.time()
  QR_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("2by2 matrix QR own")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
  m1 <- matrix( c ( 1, 2, 3, 4 ), ncol=2)
  t0 = Sys.time()
  QR(m1)
  time_lu = Sys.time() - t0
  
  sum1 = sum1 + time_lu
}
print("2by2 matrix QR")
print(sum1/1000)
```

#timing factorization methods on the 3x3 matrices
```{r}
sum2 = 0
for (i in range(1000)) {
  m1 <- matrix( c ( 1, 2, 3, 4, 5, 6, 7, 8, 9 ), nrow=3, byrow=TRUE)
  t0 = Sys.time()
  lu.decomposition(m1)
  time_lu = Sys.time() - t0
  sum2 = sum2 + time_lu
}
print("3by3 matrix LU own")
print(sum2/1000)

sum2 = 0
for (i in range(1000)) {
  m1 <- matrix( c ( 1, 2, 3, 4, 5, 6, 7, 8, 9 ), nrow=3, byrow=TRUE)
  t0 = Sys.time()
  LU_own(m1)
  time_lu = Sys.time() - t0
  sum2 = sum2 + time_lu
}
print("3by3 matrix LU")
print(sum2/1000)

sum2 = 0
for (i in range(1000)) {
  m1 <- matrix( c ( 1, 2, 3, 4, 5, 6, 7, 8, 9 ), ncol=3)
  t0 = Sys.time()
  QR(m1)
  time_qr = Sys.time() - t0
  sum2 = sum2 + time_qr
}
print("3by3 matrix QR")
print(sum2/1000)

sum2 = 0
for (i in range(1000)) {
  m1 <- matrix( c ( 1, 2, 3, 4, 5, 6, 7, 8, 9 ), ncol=3)
  t0 = Sys.time()
  QR_own(m1)
  time_qr = Sys.time() - t0
  sum2 = sum2 + time_qr
}
print("3by3 matrix QR own")
print(sum2/1000)
```

#timing factorization methods on the 4x4 matrices
```{r}
sum3 = 0
for (i in range(1000)) {
  m1 <- matrix( c (23, 34, 5, 6, 7, 12, 23, 45,  56, 76, 89, 89, 12, 3, 59, 93), nrow=4, byrow=TRUE)
  t0 = Sys.time()
  lu.decomposition(m1)
  time_lu = Sys.time() - t0
  sum3 = sum3 + time_lu
}
print("4by4 matrix LU own")
print(sum3/1000)

sum3 = 0
for (i in range(1000)) {
  m1 <- matrix( c (23, 34, 5, 6, 7, 12, 23, 45,  56, 76, 89, 89, 12, 3, 59, 93), nrow=4, byrow=TRUE)
  t0 = Sys.time()
  LU_own(m1)
  time_lu = Sys.time() - t0
  sum3 = sum3 + time_lu
}
print("4by4 matrix LU")
print(sum3/1000)

sum3 = 0
for (i in range(1000)) {
  m1 <- matrix( c (23, 34, 5, 6, 7, 12, 23, 45,  56, 76, 89, 89, 12, 3, 59, 93), ncol=4)
  t0 = Sys.time()
  QR(m1)
  time_lu = Sys.time() - t0
  sum3 = sum3 + time_lu
}
print("4by4 matrix QR")
print(sum3/1000)

sum3 = 0
for (i in range(1000)) {
  m1 <- matrix( c (23, 34, 5, 6, 7, 12, 23, 45,  56, 76, 89, 89, 12, 3, 59, 93), ncol=4)
  t0 = Sys.time()
  QR_own(m1)
  time_lu = Sys.time() - t0
  sum3 = sum3 + time_lu
}
print("4by4 matrix QR own")
print(sum3/1000)
```

#timing factorization methods on the 5x5 matrices
```{r}
sum4 = 0
for (i in range(1000)) {
   m1 <- matrix( c (23, 34, 5, 6, 7, 12, 23, 45,  56, 76, 89, 89, 12, 3, 59, 93, 2, 45, 6, 5, 11, 7, 87, 56, 4), nrow=5, byrow=TRUE)
  t0 = Sys.time()
  lu.decomposition(m1)
  time_lu = Sys.time() - t0
  sum4 = sum4 + time_lu
}
print("5by5 matrix LU own")
print(sum4/1000)

sum4 = 0
for (i in range(1000)) {
  m1 <- matrix( c (23, 34, 5, 6, 7, 12, 23, 45,  56, 76, 89, 89, 12, 3, 59, 93, 2, 45, 6, 5, 11, 7, 87, 56, 4), nrow=5, byrow=TRUE)
  t0 = Sys.time()
  LU_own(m1)
  time_lu = Sys.time() - t0
  sum4 = sum4 + time_lu
}
print("5by5 matrix LU")
print(sum4/1000)

sum4 = 0
for (i in range(1000)) {
   m1 <- matrix( c (23, 34, 5, 6, 7, 12, 23, 45,  56, 76, 89, 89, 12, 3, 59, 93, 2, 45, 6, 5, 11, 7, 87, 56, 4), ncol = 5)
  t0 = Sys.time()
  QR(m1)
  time_lu = Sys.time() - t0
  sum4 = sum4 + time_lu
}
print("5by5 matrix QR")
print(sum4/1000)

sum4 = 0
for (i in range(1000)) {
   m1 <- matrix( c (23, 34, 5, 6, 7, 12, 23, 45,  56, 76, 89, 89, 12, 3, 59, 93, 2, 45, 6, 5, 11, 7, 87, 56, 4), ncol = 5)
  t0 = Sys.time()
  QR_own(m1)
  time_lu = Sys.time() - t0
  sum4 = sum4 + time_lu
}
print("5by5 matrix QR own")
print(sum4/1000)
```

#timing factorization methods on the 6x6 matrices
```{r}
sum5 = 0
for (i in range(1000)) {
 m1 <- matrix( c (35, 67, 876, 456, 21, 11, 45, 90, 678, 0, 16, 65, 7890, 2345, 12, 5, 6, 8, 234, 567, 43, 34, 67889, 12, 56, 456, 6789, 12, 45, 67788, 0, 10, 56, 1000, 3, 1), nrow=6, byrow=TRUE)
  t0 = Sys.time()
  lu.decomposition(m1)
  time_lu = Sys.time() - t0
  sum5 = sum5 + time_lu
}
print("6by6 matrix LU own")
print(sum5/1000)

sum5 = 0
for (i in range(1000)) {
 m1 <- matrix( c (35, 67, 876, 456, 21, 11, 45, 90, 678, 0, 16, 65, 7890, 2345, 12, 5, 6, 8, 234, 567, 43, 34, 67889, 12, 56, 456, 6789, 12, 45, 67788, 0, 10, 56, 1000, 3, 1), nrow=6, byrow=TRUE)
  t0 = Sys.time()
  lu.decomposition(m1)
  time_lu = Sys.time() - t0
  sum5 = sum5 + time_lu
}
print("6by6 matrix LU")
print(sum5/1000)

sum5 = 0
for (i in range(1000)) {
 m1 <- matrix( c (35, 67, 876, 456, 21, 11, 45, 90, 678, 0, 16, 65, 7890, 2345, 12, 5, 6, 8, 234, 567, 43, 34, 67889, 12, 56, 456, 6789, 12, 45, 67788, 0, 10, 56, 1000, 3, 1), ncol=6)
  t0 = Sys.time()
  QR(m1)
  time_lu = Sys.time() - t0
  sum5 = sum5 + time_lu
}
print("6by6 matrix QR")
print(sum5/1000)

sum5 = 0
for (i in range(1000)) {
 m1 <- matrix( c (35, 67, 876, 456, 21, 11, 45, 90, 678, 0, 16, 65, 7890, 2345, 12, 5, 6, 8, 234, 567, 43, 34, 67889, 12, 56, 456, 6789, 12, 45, 67788, 0, 10, 56, 1000, 3, 1), ncol=6)
  t0 = Sys.time()
  QR_own(m1)
  time_lu = Sys.time() - t0
  sum5 = sum5 + time_lu
}
print("6by6 matrix QR own")
print(sum5/1000)
```

```{r}
big_m <- matrix(c(5, 6, 8, 10, 16, 21, 25, 37, 40, 41, 44, 52, 60, 65, 66, 67, 71, 72, 74, 76, 79, 81, 85, 88, 89, 90, 91, 94, 99, 100, 102, 104, 106, 113, 116, 120, 134, 138, 154, 172, 173, 175, 178, 184, 191, 203, 207, 210, 211, 212, 213, 215, 219, 220, 222, 226, 227, 235, 236, 238, 240, 250, 253, 262, 263, 273, 274, 280, 281, 291, 293, 304, 305, 307, 310, 312, 315, 317, 318, 321, 326, 330, 334, 335, 336, 340, 341, 344, 349, 362, 364, 367, 373, 374, 375, 376, 377, 380, 381, 390), nrow = 10, byrow = TRUE)
```

#timing factorization methods on the 10by10
```{r}
sum1 = 0
for (i in range(1000)) {
  t0 = Sys.time()
  lu.decomposition(big_m)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("10by10 matrix LU")
print(sum1/1000)

sum5 = 0
for (i in range(1000)) {
  t0 = Sys.time()
  LU_own(big_m)
  time_lu = Sys.time() - t0
  sum5 = sum5 + time_lu
}
print("10by10matrix LU own")
print(sum5/1000)

sum2 = 0
for (i in range(1000)) {
  t0 = Sys.time()
  QR(big_m)
  time_lu = Sys.time() - t0
  sum2 = sum2 + time_lu
}
print("10by10 matrix QR")
print(sum2/1000)

sum4 = 0
for (i in range(1000)) {
  t0 = Sys.time()
  QR_own(big_m)
  time_lu = Sys.time() - t0
  sum4 = sum4 + time_lu
}
print("10by10 matrix QR own")
print(sum4/1000)
```

```{r}
big_m_30 <- matrix(c(611, 342, 594, 136, 701, 498, 555, 415, 299, 331, 346, 632, 507, 509, 192, 24, 237, 99, 56, 687, 565, 879, 728, 262, 141, 812, 430, 175, 561, 527, 474, 86, 254, 250, 831, 779, 365, 356, 659, 255, 458, 520, 406, 475, 278, 150, 792, 544, 407, 76, 522, 434, 364, 759, 823, 834, 265, 629, 501, 609, 155, 644, 706, 638, 766, 688, 138, 705, 884, 352, 161, 843, 763, 404, 764, 435, 826, 658, 85, 280, 755, 191, 882, 564, 199, 895, 648, 459, 429, 657, 615, 877, 400, 335, 736, 530, 827, 245, 476, 771, 672, 313, 127, 676, 623, 295, 722, 140, 93, 326, 239, 101, 306, 298, 511, 332, 294, 3, 15, 409, 65, 487, 144, 824, 880, 859, 167, 704, 431, 207, 646, 143, 104, 412, 626, 765, 563, 508, 27, 348, 510, 622, 30, 401, 851, 51, 177, 692, 26, 118, 690, 897, 551, 419, 852, 583, 96, 117, 457, 133, 186, 388, 862, 750, 624, 389, 122, 708, 293, 21, 41, 888, 803, 221, 865, 258, 495, 319, 870, 148, 215, 296, 361, 158, 465, 32, 288, 504, 54, 585, 4, 491, 529, 485, 88, 433, 618, 890, 374, 733, 180, 443, 305, 102, 146, 566, 630, 878, 5, 860, 794, 427, 314, 188, 44, 375, 666, 203, 275, 797, 744, 773, 738, 422, 702, 714, 678, 795, 516, 205, 209, 680, 60, 243, 540, 81, 337, 494, 639, 360, 421, 679, 735, 593, 266, 114, 384, 767, 512, 703, 185, 653, 790, 484, 693, 340, 181, 586, 69, 742, 757, 715, 642, 549, 378, 513, 537, 169, 665, 277, 819, 732, 866, 34, 667, 614, 109, 730, 490, 226, 896, 747, 229, 669, 550, 548, 166, 256, 263, 440, 396, 323, 224, 31, 383, 198, 842, 640, 452, 605, 59, 606, 408, 579, 674, 184, 307, 260, 506, 110, 28, 660, 874, 359, 303, 503, 528, 82, 129, 811, 8, 753, 599, 619, 204, 524, 841, 350, 448, 193, 636, 67, 358, 43, 612, 492, 351, 840, 318, 631, 126, 157, 438, 95, 151, 799, 338, 240, 543, 137, 125, 699, 752, 602, 734, 791, 171, 402, 577, 194, 463, 464, 211, 820, 142, 894, 637, 545, 11, 64, 108, 816, 697, 225, 370, 381, 783, 284, 398, 726, 634, 534, 597, 832, 274, 271, 785, 743, 650, 116, 290, 173, 257, 425, 219, 478, 885, 179, 881, 821, 558, 651, 14, 489, 1, 709, 861, 806, 681, 53, 354, 892, 79, 380, 168, 707, 854, 214, 426, 574, 200, 283, 589, 789, 393, 369, 6, 42, 571, 853, 462, 473, 600, 445, 235, 500, 749, 695, 63, 357, 661, 775, 610, 311, 467, 353, 781, 37, 655, 247, 46, 883, 793, 829, 718, 499, 105, 547, 553, 174, 737, 55, 620, 748, 339, 470, 220, 469, 616, 713, 128, 377, 786, 486, 217, 120, 450, 36, 663, 837, 596, 802, 70, 344, 17, 441, 423, 582, 304, 645, 519, 784, 71, 22, 98, 269, 770, 115, 49, 437, 268, 635, 297, 858, 454, 446, 633, 234, 756, 830, 197, 533, 231, 846, 698, 621, 601, 568, 152, 844, 717, 38, 270, 893, 857, 113, 461, 367, 312, 213, 13, 91, 617, 131, 598, 362, 817, 560, 595, 327, 249, 696, 424, 273, 145, 447, 536, 52, 292, 12, 814, 160, 451, 301, 382, 689, 746, 647, 107, 664, 455, 436, 139, 712, 453, 246, 839, 628, 267, 684, 403, 134, 652, 170, 222, 526, 796, 867, 801, 808, 782, 518, 238, 567, 531, 588, 428, 523, 809, 871, 471, 162, 343, 800, 176, 887, 716, 259, 662, 559, 248, 347, 552, 341, 444, 591, 607, 83, 50, 285, 228, 554, 334, 212, 711, 159, 825, 562, 372, 758, 869, 710, 132, 875, 253, 838, 385, 57, 493, 873, 77, 442, 355, 649, 324, 315, 291, 580, 119, 514, 410, 891, 73, 673, 720, 772, 325, 308, 164, 135, 23, 761, 89, 391, 863, 336, 16, 279, 242, 432, 542, 103, 694, 74, 172,
576, 729, 252, 149, 876, 153, 849, 538, 604, 686, 656, 575, 309, 818, 287, 182, 724, 505, 154, 833, 727, 439, 572, 480, 90, 48, 272, 368, 815, 677, 768, 92, 592, 405, 804, 395, 227, 80, 147, 100, 627, 201, 741, 189, 670, 97, 625, 206, 130, 261, 223, 232, 456, 9, 2, 899, 330, 745, 289, 805, 282, 449, 836, 872, 613, 780, 886, 84, 810, 366, 392, 845, 264, 413, 850, 578, 379, 788, 251, 345, 488, 300, 671, 725, 420, 541, 373, 822, 898, 691, 557, 25, 546, 390, 376, 19, 208, 66, 483, 47, 723, 317, 45, 394, 654, 481, 889, 517, 502, 641, 774, 776, 230, 835, 363, 590, 320, 163, 570, 187, 587, 124, 29, 497, 416, 798, 608, 371, 569, 479, 418, 316, 813, 399, 322, 10, 18, 178, 477, 183, 643, 112, 302, 397, 535, 276, 856, 521, 532, 864, 496, 754, 855, 466, 20, 828, 216, 75, 210, 581, 482, 668, 333, 900, 525, 515, 848, 244, 414, 94, 68, 62, 78, 165, 556, 190, 286, 195, 58, 847, 584, 196, 472, 329, 751, 281, 35, 241, 675, 33, 40, 7, 387, 321, 740, 807, 739, 39, 121, 417, 787, 700, 685, 218, 156, 310, 123, 61, 682, 760, 719, 233, 111, 539, 72, 106, 762, 778, 731, 573, 769, 603, 411, 683, 328, 721, 468, 202, 87, 236, 349, 777, 460, 868, 386), nrow = 30, byrow = TRUE)
```

#30by30 timing
```{r}
sum1 = 0
for (i in range(1000)) {
  t0 = Sys.time()
  lu.decomposition(big_m_30)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("30by30 matrix LU")
print(sum1/1000)

sum5 = 0
for (i in range(1000)) {
  t0 = Sys.time()
  LU_own(big_m_30)
  time_lu = Sys.time() - t0
  sum5 = sum5 + time_lu
}
print("30by30matrix LU own")
print(sum5/1000)

sum2 = 0
for (i in range(1000)) {
  t0 = Sys.time()
  QR(big_m_30)
  time_lu = Sys.time() - t0
  sum2 = sum2 + time_lu
}
print("30by30 matrix QR")
print(sum2/1000)

sum4 = 0
for (i in range(1000)) {
  t0 = Sys.time()
  QR_own(big_m_30)
  time_lu = Sys.time() - t0
  sum4 = sum4 + time_lu
}
print("30by30 matrix QR own")
print(sum4/1000)
```

Part that includes time tests on not only square matrices for QR factorization.
#timing factorization methods on the 3x6 matrices
```{r}
sum1 = 0
for (i in range(1000)) {
 m1 <- matrix(c(3, 2, 3, 2, -1, 4, 5, 1, -5, 4, 2, 1, 9, -3, 2 , -1, 8, 1), nrow=3, byrow=TRUE)
  t0 = Sys.time()
  QR(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by6 matrix QR")
print(sum1/1000)

sum2 = 0
for (i in range(1000)) {
 m1 <- matrix(c(3, 2, 3, 2, -1, 4, 5, 1, -5, 4, 2, 1, 9, -3, 2 , -1, 8, 1), nrow=3, byrow=TRUE)
  t0 = Sys.time()
  QR_own(m1)
  time_lu = Sys.time() - t0
  sum2 = sum2 + time_lu
}
print("3by6 matrix QR own")
print(sum2/1000)
```

#timing factorization methods on the 3x4 matrices
```{r}
sum1 = 0
for (i in range(1000)) {
 m1 <- matrix(c(1, -1, -1, 1, 0, 0, 1, -1, 0, 0, 0, -1), nrow=3, byrow=TRUE)
  t0 = Sys.time()
  QR(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("3by4 matrix QR")
print(sum1/1000)

sum2 = 0
for (i in range(1000)) {
 m1 <- matrix(c(1, -1, -1, 1, 0, 0, 1, -1, 0, 0, 0, -1), nrow=3, byrow=TRUE)
  t0 = Sys.time()
  QR_own(m1)
  time_lu = Sys.time() - t0
  sum2 = sum2 + time_lu
}
print("3by4 matrix QR own")
print(sum2/1000)
```
#timing factorization methods on the 4x5 matrices
```{r}
sum1 = 0
for (i in range(1000)) {
 m1 <- matrix(c(3, 2, 3, 2, -1, 4, 5, 1, -5, 4, 2, 1, 9, -3, 2 , -1, 10, -11, 0 , 2), nrow=4, byrow=TRUE)
  t0 = Sys.time()
  QR(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("4by5 matrix QR")
print(sum1/1000)

sum1 = 0
for (i in range(1000)) {
 m1 <- matrix(c(3, 2, 3, 2, -1, 4, 5, 1, -5, 4, 2, 1, 9, -3, 2 , -1, 10, -11, 0 , 2), nrow=4, byrow=TRUE)
  t0 = Sys.time()
  QR_own(m1)
  time_lu = Sys.time() - t0
  sum1 = sum1 + time_lu
}
print("4by5 matrix QR own")
print(sum1/1000)
```